---
title: "Data Tidying"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Best practice is to put all library calls in same R chunk and up front.

The order in which you order your libraries matters.

```{r message = FALSE,warning=FALSE}
library(dplyr)
library(tidyr)
library(readr)
```

Other functions you may use:

- package_name::function eg dplyr::filter 
- install.packages("dplyr")

## Keyboard shortcuts 

R code chunk:
- cmd + option + i (Mac)
- ctrl + alt + i (PC)

Pipe operator
- Ctrl + Shift + M (PC)

Hints:
- read_csv() is a bit smarter and more efficient than - read.csv and gives you a readout

```{r}
catch_original <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")
```
## Clean data

- drop All and notes cols using select (or select -colname to drop)

```{r}
catch_data <- catch_original %>% 
  # select(Region, Year, Chinook, Coho, Pink, Chum, Sockeye)
  select(-All, -notesRegCode)

head(catch_data)

summary(catch_data)
```
 - change Chinook from character to numeric type
```{r}
catch_clean <- catch_data %>% 
  mutate(Chinook = if_else(Chinook == "I", "1", Chinook)) %>% 
  mutate(Chinook = as.numeric(Chinook))

head(catch_clean)

summary(catch_clean)
```
### finding spurious characters in Chinook col
- which function returns a value for which an argument is true.

```{r}
i <- which(is.na(catch_clean$Chinook))
           i
catch_original[i, ]           
```
## Reshaping data (wide to long)
- move from wide to long format

```{r}
catch_long <- catch_clean %>% 
  pivot_longer(cols = c(-Region, -Year), names_to = "Species", values_to = "Catch")
head(catch_long)
```
```{r}
catch_wide <- catch_long %>% 
  pivot_wider(names_from = Region, values_from = Catch)

head(catch_wide)
```

Hints:
- If column names exist (as in names_from and values_from), call them directly (without quotes) as variables.
- If column names are new (as in names_to and values_to), you need to call them with quotes since you are creating them
- Can use rename function (takes same args as mutate() function) to rename col heads

```{r}
catch_long <- catch_clean %>% 
  pivot_longer(cols = c(-Region, -Year), names_to = "Species", values_to = "Catch") %>% 
  rename(catch_thousands = Catch)

catch_long <- catch_long %>% 
  mutate(catch = catch_thousands * 1000) %>% 
  select(-catch_thousands) #hide catch_thousands column

head(catch_long)
```


## Split-Apply-Combine

- mean catch by region

```{r}
mean_region <- catch_long %>% 
  group_by(Region, Species) %>% 
  summarise(catch_mean = mean(catch), n=n(), .groups = "drop") #try also without .groups = "drop"

head(mean_region)
```
 - add a filter argument (same as subset in tidyverse)

```{r}
mean_region <- catch_long %>% 
  filter(Species == "Sockeye") %>% 
  group_by(Region, Species) %>% 
  summarise(catch_mean = mean(catch), n=n(), .groups = "drop") #try also without .groups = "drop"

head(mean_region)
```
### Joins

- implement a left join
- often a necessary step, esp. with normalized data

```{r}
region_defs <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1") %>% 
    select(code, mgmtArea)

head(region_defs)
```

```{r}
catch_joined <- left_join(catch_long, region_defs, by = c("Region" = "code")) #because no common col names you have to include "by" arg and put col names in "" here. These are also case specific.

head(catch_joined)
```
- to check if number of rows is the same in both

```{r}
catch_joined <- left_join(catch_long, region_defs, by = c("Region" = "code")) 

nrow(catch_joined) == nrow(catch_long) #should return TRUE (it does)
```

