---
title: "Untitled"
author: "Jana Peirce"
date: "2/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = FALSE, message = FALSE}
library(sf)
library(ggplot2)
library(leaflet)
library(scales)
library(ggmap)
library(dplyr)
```

```{r}
ak_regions <- read_sf("../shapefiles/ak_regions_simp.shp")

plot(ak_regions) 
```
```{r}
class(ak_regions)
```

```{r}
head(ak_regions)
```
```{r}
ak_regions_3338 <- ak_regions %>% #use pipe because SF is part of tidyverse/dplyr
  st_transform(crs = 3338)

st_crs(ak_regions_3338)#this is a good CRS projection for AK
```

```{r}
plot(ak_regions_3338)
```
```{r}
ak_regions_3338 %>%
  select(region)
```

```{r}
pop <- read.csv("../shapefiles/alaska_population.csv")

head(pop)
```

```{r}
class(pop)
```
- Do a spatial join using st_join() - look at documentation for st_join for other joins
```{r}
pop_4326 <- st_as_sf(pop, 
                  coords = c('lng', 'lat'),
                  crs = 4326, #4236 is WGS84. You have to read in the data in the format it's in and then transform it. You cannot just assign it a diff CRS (such as 3338)
                  remove = F)

head(pop_4326)
class(pop_4326)
```
- Oops an error! Because they use differnt projections

```{r}
pop_joined <- st_join(pop_4326, ak_regions_3338, join = st_within)
```

```{r}
pop_3338 <- st_transform(pop_4326, crs = 3338)

head(pop_3338)
```
```{r}
pop_joined <- st_join(pop_3338, ak_regions_3338, join = st_within)

head(pop_joined)
```
- we can drop geometries now, since we no longer need it. But since GEOMETRIES are sticky you need to unstick it using "as.data.frame" to convert if from SF table to normal dataframe

```{r}
pop_region <- pop_joined %>% 
  as.data.frame() %>% 
  group_by(region) %>% 
  summarise(total_pop = sum(population))

head(pop_region)
```

- Do a regular left_join to add the total_pop to the original map talbe

```{r}
pop_region_3338 <- left_join(ak_regions_3338, pop_region)

#joining by region

plot(pop_region_3338)
```
```{r}
pop_mgmt_3338 <- pop_region_3338 %>% 
  group_by(mgmt_area) %>% 
  summarize(total_pop = sum(total_pop), do_union = F) #do union to get rid of holes???

head(pop_mgmt_3338)
plot(pop_mgmt_3338["mgmt_area"])
```

- map with ggplot. include args explicitly instead of shorthand, bc geom_sf puts data and mapping args in reverse order!

```{r}
ggplot(pop_region_3338) +
  geom_sf(aes(fill = total_pop)) 

```
 
- and with some refining...

```{r}
ggplot(pop_region_3338) +
  geom_sf(aes(fill = total_pop)) +
  theme_bw() +
  labs(fill = "Total Population") +
  scale_fill_continuous(low = "khaki", high =  "firebrick", labels = comma)
```

- Need to add point and river data

```{r}
rivers_3338 <- read_sf("../shapefiles/ak_rivers_simp.shp")
st_crs(rivers_3338)

head(rivers_3338)
```

```{r}
ggplot(pop_region_3338) +
  geom_sf(aes(fill = total_pop)) +
  geom_sf(data = rivers_3338, mapping = aes(size = StrOrder)) +
  theme_bw() +
  labs(fill = "Total Population") +
  scale_fill_continuous(low = "khaki", high =  "firebrick", labels = comma)
```

- need to scale size of rivers to something skinner - guess and check!

```{r}
ggplot(pop_region_3338) +
  geom_sf(aes(fill = total_pop)) +
  geom_sf(data = rivers_3338, mapping = aes(size = StrOrder)) +
  geom_sf(data = pop_3338, mapping = aes(), size = 0.5) +
  scale_size(range = c(0.01, 0.2), guide = "none") +
  theme_bw() +
  labs(fill = "Total Population") +
  scale_fill_continuous(low = "khaki", high =  "firebrick", labels = comma)
```

- package ggmap let's you plot your own points on a basemap (base tiles)
- unlike leaflet you can't zoom around on it.
- basemap is in CRS 3857, so we need to transform our data to match

```{r}
pop_3857 <- pop_3338 %>%
  st_transform(crs = 3857)
```

- get the basemap, but first fix an issue with ggmap with this function

```{r}
# Define a function to fix the bbox to be in EPSG:3857
# See https://github.com/dkahle/ggmap/issues/160#issuecomment-397055208
ggmap_bbox_to_3857 <- function(map) {
  if (!inherits(map, "ggmap")) stop("map must be a ggmap object")
  # Extract the bounding box (in lat/lon) from the ggmap to a numeric vector, 
  # and set the names to what sf::st_bbox expects:
  map_bbox <- setNames(unlist(attr(map, "bb")), 
                       c("ymin", "xmin", "ymax", "xmax"))
  
  # Coonvert the bbox to an sf polygon, transform it to 3857, 
  # and convert back to a bbox (convoluted, but it works)
  bbox_3857 <- st_bbox(st_transform(st_as_sfc(st_bbox(map_bbox, crs = 4326)), 3857))
  
  # Overwrite the bbox of the ggmap object with the transformed coordinates 
  attr(map, "bb")$ll.lat <- bbox_3857["ymin"]
  attr(map, "bb")$ll.lon <- bbox_3857["xmin"]
  attr(map, "bb")$ur.lat <- bbox_3857["ymax"]
  attr(map, "bb")$ur.lon <- bbox_3857["xmax"]
  map
}
```

- basemap we are using is stamenmap

```{r}
bbox <- c(-170, 52, -130, 68)   # This is roughly southern Alaska
ak_map <- get_stamenmap(bbox, zoom = 4)
ak_map_3857 <- ggmap_bbox_to_3857(ak_map)
```

- use inherit.aes = FALSE because there is stuff going on in the background

```{r}
ggmap(ak_map_3857) + 
  geom_sf(data = pop_3857, aes(color = population), inherit.aes = F) +
  scale_color_continuous(low = "khaki", high =  "firebrick", labels = comma)
```

- there is a geom_raster that you can use if you have a raster basemap you want to use.

# Visualize SF objects with leaflet

- pasted from book but not running or generating error message

```{r}
epsg3338 <- leaflet::leafletCRS(
  crsClass = "L.Proj.CRS",
  code = "EPSG:3338",
  proj4def =  "+proj=aea +lat_1=55 +lat_2=65 +lat_0=50 +lon_0=-154 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs",
  resolutions = 2^(16:7))
```

```{r}
st_crs(pop_region_3338)
```

```{r}
pop_region_4326 <- pop_region_3338 %>% st_transform(crs = 4326)
```

```{r}
m <- leaflet(options = leafletOptions(crs = epsg3338)) %>%
        addPolygons(data = pop_region_4326, 
                    fillColor = "gray",
                    weight = 1)

m
```

- We can add labels, legends, and a color scale.

```{r}
pal <- colorNumeric(palette = "Reds", domain = pop_region_4326$total_pop)

m <- leaflet(options = leafletOptions(crs = epsg3338)) %>%
        addPolygons(data = pop_region_4326, 
                    fillColor = ~pal(total_pop),
                    weight = 1,
                    color = "black",
                    fillOpacity = 1,
                    label = ~region) %>% 
        addLegend(position = "bottomleft",
                  pal = pal,
                  values = range(pop_region_4326$total_pop),
                  title = "Total Population")

m
```

- We can also add the individual communities, with popup labels showing their population, on top of that!

```{r}
pal <- colorNumeric(palette = "Reds", domain = pop_region_4326$total_pop)

m <- leaflet(options = leafletOptions(crs = epsg3338)) %>%
        addPolygons(data = pop_region_4326, 
                    fillColor = ~pal(total_pop),
                    weight = 1,
                    color = "black",
                    fillOpacity = 1) %>% 
        addCircleMarkers(data = pop_4326,
                         lat = ~lat,
                         lng = ~lng,
                         radius = ~log(population/500), # arbitrary scaling
                         fillColor = "gray",
                         fillOpacity = 1,
                         weight = 0.25,
                         color = "black",
                         label = ~paste0(pop_4326$city, ", population ", comma(pop_4326$population))) %>%
        addLegend(position = "bottomleft",
                  pal = pal,
                  values = range(pop_region_4326$total_pop),
                  title = "Total Population")

m
```

- There is a lot more functionality to sf including the ability to intersect polygons, calculate distance, create a buffer, and more. Here are some more great resources and tutorials for a deeper dive into this great package: << See book for links! >>

