---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Header 1

## Header 2

### Header 3

Write some **bold text**

-   list item
-   list item
-   list item

This is *italic text*

1.  Item 1

2.  Item 2

3.  Item 3

    -   Item 3a
    -   Item 3b

### Help Docs

Help > Markdown Quick Reference
Help > Cheet Sheets > Markdown Cheat Sheet (more compact PDF format)
    
### Shortcuts for creating code chunk and     

Mac shortcut for R code chunk: cmd + option + i
PC shortcut: Ctrl + Alt + i

```{r}
x <- 3 + 2
```

### Shortcuts for running code chunk 

Running an R chunk:
Mac: 
PC: Ctrl + enter (can run individual expressions - 1 line at a time)
If you use Play button or Run menu it runs the whole code chunk 

```{r}
y <- 5 - 3
y + 10
```
## Read in data

1. create data folder
1. upload data file
1. create R chunk as below and type in read.csv
1. use tabs to see autofill options

```{r}
bg_chem <- read.csv("../data/BGchem2008data.csv")
```




## Analysis

## Plots

```{r, echo=FALSE}
plot(bg_chem$CTD_Depth, bg_chem$CTD_Temperature)
```
 
## Git and GitHUB

I learned that you cannot use copy and paste in a Windows terminal window.