---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Header 1

## Header 2

### Header 3

Write some **bold text**

-   list item
-   list item
-   list item

This is *italic text*

1.  Item 1

2.  Item 2

3.  Item 3

    -   Item 3a
    -   Item 3b

### Help Docs

Help > Markdown Quick Reference
Help > Cheet Sheets > Markdown Cheat Sheet (more compact PDF format)
    
### Shortcuts for creating code chunk and     

Mac shortcut for R code chunk: cmd + option + i
PC shortcut: Ctrl + Alt + i

```{r}
x <- 3 + 2
```

### Shortcuts for running code chunk 

Running an R chunk:
Mac: 
PC: Ctrl + enter (can run individual expressions - 1 line at a time)
If you use Play button or Run menu it runs the whole code chunk 

```{r}
y <- 5 - 3
y + 10
```
## Read in data

1. create data folder
1. upload data file
1. create R chunk as below and type in read.csv
1. use tabs to see autofill options

```{r}
bg_chem <- read.csv("../data/BGchem2008data.csv")
```




## Analysis

## Plots

```{r, echo=FALSE}
plot(bg_chem$CTD_Depth, bg_chem$CTD_Temperature)
```
 
## Git and GitHUB

I learned that you cannot use copy and paste keyboard commands in a Windows terminal window. 

You can right-click and choose paste or copy from the drop-down menu.

## Archiving with GitHub

- GitHub is not an archive.
- Software Heritage - is an archive and makes it easty to archive from GitHub. Most relevant for software processes that will be useful for others independent of your data and R environment.
- Arctic Data Center another place to archive data and code

## Collaborating in GitHub repositories without Conflicts 5.3

See appendix for info on sharing with "Pull Requests"
For trusted colleagues can use direct sharing.

Owner: in GitHub > Settings > Collaborators > Add People (know their username)

Collaborator: Will receive email with link. 

In GitHub: Go to Code button dropdown menu and under clone, choose copy

In Rstudio: Create new project with Git

If you make changes

PULL before Commit (always wise)
If there are any conflicts, it will let you know that. You can grab other changes and then merge your changes in the doc before you push your version,

Can do "Git Fetch" from CMD line instead to see copies of both files.

Owner: Pull any changes whenever opening up project in RStudio again.

Git commands to resolve conflicts:
- git merge --abort
- git checkout --theirs conflicted_file.Rmd
- git checkout --ours conflicted_file.Rmd
Then run "add, "commit" and "push" as normal.